Manifold Clustering in the Setting of Generalized Random Dot Product Graphs

### 2023-05-24 sdss-2023/slides.Rmd
updated slides

gridExtra::grid.arrange(latent.vecs, rdpg, ase, nrow = 1)
```

@@

>>>>

# gridExtra::grid.arrange(latent.vecs, rdpg, ase, nrow = 1)
gridExtra::grid.arrange(latent.vecs, rdpg, nrow = 1)
```

@@

## Asymptotic Results

**Theorem**. 
Let an MBM be such that the manifolds are described by functions $g_1(t), ..., g_K(t)$ which are polynomial curves of order $R$. 
Define the loss of $K$-curves clustering as:

$$L(\hat{z}_1, ..., \hat{z}_n, \hat{g}_1, ..., \hat{g}_K; A) = \sum_k \sum_{i : \hat{z}_i = k} \|\hat{x}_{i} - \hat{g}_k(t_{i})\|^2,$$

where $\hat{x}_i$ are the embedding vectors of $A$. 
Suppose that for each community $k$, we have labels for at least $R + 1$ vertices. 
Then as $n \to \infty$, $K$-curves clustering outputs estimates such that

$$L(\hat{z}_1, ..., \hat{z}_n, \hat{g}_1, ..., \hat{g}_K; A) \stackrel{p}{\to} 0.$$

## Simulation

```{r three-curves, cache = TRUE, fig.height = 2, fig.width = 6, out.width = '100%'}
set.seed(314159)
par(mfrow = c(1, 3))

n <- 256

p.list <- list(matrix(c(0, 1, 0, 
                        0, 0, 1,
                        0, 0, 0),
                      nrow = 3, ncol = 3),
               matrix(c(0, 0, 0, 
                        0, 0, 1,
                        0, 1, 0),
                      nrow = 3, ncol = 3),
               matrix(c(0, 1, 0, 
                        0, 0, 1,
                        0, 1, 0),
                      nrow = 3, ncol = 3))

K <- length(p.list)

z <- sample(seq(K), n, replace = TRUE)
z <- sort(z)

t <- runif(n)

X <- lapply(seq(K), function(k) {
  bezier.curve(t[z == k], 
               p.list[[k]], 
               intercept = TRUE)
}) %>% 
  do.call('rbind', .)

p <- ncol(X)
q <- 0

P <- grdpg.edge.prob.matrix(X, p, q)
A <- draw.graph(P)
Xhat <- embedding(A, p, q)

degs <- colSums(A)
A <- A[degs > 0, degs > 0]
z <- z[degs > 0]
X <- X[degs > 0, ]
Xhat <- Xhat[degs > 0, ]

scatterplot3d::scatterplot3d(X[, 1], X[, 2], X[, 3],
                             color = z + 1, 
                             mar = c(0, 0, 0, 0),
                             xlab = NA, ylab = NA, zlab = NA,
                             # pch = '.', 
                             asp = 1,
                             tick.marks = FALSE)

qgraph::qgraph(A, 
               groups = factor(z), 
               borders = FALSE,
               legend = FALSE,
               esize = .001,
               mar = c(0, 0, 0, 0))

scatterplot3d::scatterplot3d(Xhat[, 1], Xhat[, 2], Xhat[, 3], 
                             color = z + 1, 
                             mar = c(0, 0, 0, 0),
                             xlab = NA, ylab = NA, zlab = NA, 
                             # pch = '.', 
                             asp = 1,
                             # angle = 40,
                             tick.marks = FALSE)
```

```{r sim-curves-3, fig.height = 2.5, fig.width = 6, out.width = '100%', cache = TRUE}
n.vec <- 2 ^ c(7, 8, 9, 10, 11)
nsamp.vec <- c(0, 4, 8)

clustering.df <- readr::read_csv('~/dev/manifold-block-models/simulations/balanced-3/balanced-3.csv')

clust.summary.df <- clustering.df %>% 
  dplyr::group_by(n, nsamp) %>% 
  dplyr::summarise(
    med.count = median(error.count),
    first.q.count = quantile(error.count, .25),
    third.q.count = quantile(error.count, .75),
    med.rate = median(error.rate),
    first.q.rate = quantile(error.rate, .25),
    third.q.rate = quantile(error.rate, .75)
  ) %>% 
  dplyr::ungroup()

ggplot(clust.summary.df) + 
  theme_bw() + 
  theme(text = element_text(size = 10)) + 
  scale_y_log10() +
  scale_x_log10(breaks = n.vec) +
  labs(y = 'error rate',
       x = 'n',
       colour = 'known labels per community', 
       shape = NULL) + 
  geom_line(aes(x = n, y = med.rate, colour = factor(nsamp))) + 
  geom_point(aes(x = n, y = med.rate, colour = factor(nsamp))) + 
  geom_errorbar(aes(x = n, 
                    ymin = first.q.rate, ymax = third.q.rate, 
                    colour = factor(nsamp)),
                width = .125) + 
  theme(legend.position = 'bottom') + 
  scale_colour_brewer(palette = 'Set1')
```

## Thank you!

Code and drafts available at https://github.com/johneverettkoo/manifold-block-models

### 2023-05-24 sdss-2023/slides.tex
updated slides

>>>>

\begin{frame}{Asymptotic Results}
\protect\hypertarget{asymptotic-results}{}
\textbf{Theorem}. Let an MBM be such that the manifolds are described by
functions \(g_1(t), ..., g_K(t)\) which are polynomial curves of order
\(R\). Define the loss of \(K\)-curves clustering as:

\[L(\hat{z}_1, ..., \hat{z}_n, \hat{g}_1, ..., \hat{g}_K; A) = \sum_k \sum_{i : \hat{z}_i = k} \|\hat{x}_{i} - \hat{g}_k(t_{i})\|^2,\]

where \(\hat{x}_i\) are the embedding vectors of \(A\). Suppose that for
each community \(k\), we have labels for at least \(R + 1\) vertices.
Then as \(n \to \infty\), \(K\)-curves clustering outputs estimates such
that

\[L(\hat{z}_1, ..., \hat{z}_n, \hat{g}_1, ..., \hat{g}_K; A) \stackrel{p}{\to} 0.\]
\end{frame}

\begin{frame}{Simulation}
\protect\hypertarget{simulation}{}
\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/three-curves-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{slides_files/figure-beamer/sim-curves-3-1} \end{center}
\end{frame}

\begin{frame}{Thank you!}
\protect\hypertarget{thank-you}{}
Code and drafts available at
\url{https://github.com/johneverettkoo/manifold-block-models}
\end{frame}
